---
title: "myTinyShinys"
author: "pssguy"
date: "Tuesday, 19th May 2015"
output:
  html_document:
    fig_caption: yes
resource_files:
- ne_50m_admin_0_countries.dbf
- ne_50m_admin_0_countries.shp
- ne_50m_admin_0_countries.shx
- ne_50m_admin_0_countries.prj
runtime: shiny
css: custom.css
---
###  <span class="post">Climate Change Views</span>
*19th May 2015*

I am currently dipping into a MOOC on [Climate Denial](https://www.edx.org/course/making-sense-climate-science-denial-uqx-denial101x). 
   My small contribution is to map some rather outdated (2007/8) Gallup poll data from a survey of individuals from 128 countries in the first comprehensive study of global opinions.The information is available in a [Wikipedia Article](http://en.wikipedia.org/wiki/Climate_change_opinion_by_country)


Enter the category you are interested in and the number of groupings (2-9) you wish to separate the countries into. Click on any country for all the data
```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
# load required libraries

library(XML)
library(httr)
library(stringr)
library(leaflet)
library(rgdal)
library(dplyr)
library(ggvis)

## could download and save to file which is read in but
## this shows method and is pretty brief anyways 

## data is in html table so this is easiest method to extract
theUrl <- "http://en.wikipedia.org/wiki/Climate_change_opinion_by_country"

tables = readHTMLTable(theUrl,colClasses=c("character","integer","integer","integer"),stringsAsFactors=F)

# tables are unnamed but it is second on page
data <- tables[[2]]

#rename columns
colnames(data) <- c("Country","Awareness","Caused_by_Humans","Perceived_Threat")


# read in country shape files previously downloaded from Natural Earth
#http://www.naturalearthdata.com/downloads/50m-cultural-vectors/50m-admin-0-countries-2/

countries <- readOGR(dsn=".",
                     layer = "ne_50m_admin_0_countries", 
                     encoding = "UTF-8",verbose=FALSE)

## check for differences in country names and correct
sort(setdiff(data$Country,countries$name))

data[data$Country=="Central African Republic",]$Country <- "Central African Rep."
data[data$Country=="Czech Republic",]$Country <- "Czech Rep."
data[data$Country=="Democratic Republic of the Congo",]$Country <- "Dem. Rep. Congo"
data[data$Country=="Dominican Republic",]$Country <- "Dominican Rep."
data[data$Country=="Laos",]$Country <- "Lao PDR"
data[data$Country=="Republic of the Congo",]$Country <- "Congo"
data[data$Country=="South Korea",]$Country <- "Korea"

# merge the data with the shapefiles
countries2 <- sp::merge(countries, 
                        data, 
                        by.x = "name", 
                        by.y = "Country",                    
                        sort = FALSE) 

# create popup

popUp <- paste0("<strong>", countries2$name, "</strong><br>",
                 
                "<br><strong>Awareness: </strong>", countries2$Awareness, 
                "%<br><strong>Caused by Humans: </strong>", countries2$Caused_by_Humans,
                "%<br><strong>Perceived Threat: </strong>", countries2$Perceived_Threat,"%")


# creative inputs for groupings and categories 
numericInput("bins","Number of Groups",value=4,min=2,max=9,step=1)

radioButtons(inputId="category",label="Category",choices=c("Awareness","Caused_by_Humans","Perceived_Threat"),selected="Awareness",inline=TRUE)




# create map
output$map <- renderLeaflet({
 
  if (is.null(input$category)) return()
 
  # split into equal sized bins for all 128 counties
  theBins <- input$bins
buckets <- seq(from=1, to=128,by=128/theBins)


# Create labels for legend
for (i in 1:(length(buckets)-1)) {
  if (i!=1) {
    labs <- c(labs,paste0(floor(buckets[i]),"-",floor(buckets[i+1]-1)))
  } else {
    labs<- paste0(floor(buckets[i]),"-",floor(buckets[i+1]-1))
  }
  
}
labs <- c(labs,paste0(floor(buckets[i+1]),"-128"),"NA")

# use a qualitative color range
pal <- colorQuantile("RdYlGn", domain=NULL, n = theBins,na.color="#808080")

# need to find way to 
  if(input$category=="Awareness") {
  theMap <-countries2 %>% 
  leaflet() %>% 
  setView(lng=0,lat=0,zoom= 1) %>% 
  addTiles()  %>% # defaults to openstreetmap version
  addPolygons(fillColor = ~pal(Awareness), weight=1,color = "#BDBDC3"
              ,fillOpacity=0.9 ,
             popup = popUp) %>% 

  addLegend(colors = c(rev(RColorBrewer::brewer.pal(theBins, "RdYlGn")), "#808080"),  
            bins = theBins, 
            position = 'bottomleft', 
            title = "Ranking Band", 
            labels = labs) %>% 
  mapOptions(zoomToLimits="first") 
} else if(input$category=="Caused_by_Humans") {
 theMap <-countries2 %>% 
  leaflet() %>% 
  setView(lng=0,lat=0,zoom= 1) %>% 
  addTiles()  %>% 
  addPolygons(fillColor = ~pal(Caused_by_Humans), weight=1,color = "#BDBDC3"
              ,fillOpacity=0.9 ,
             popup = popUp) %>% 

  addLegend(colors = c(rev(RColorBrewer::brewer.pal(theBins, "RdYlGn")), "#808080"),  
            bins = theBins, 
            position = 'bottomleft', 
            title = "Ranking Band", 
            labels = labs) %>% 
  mapOptions(zoomToLimits="first") 
   } else {
      theMap <-countries2 %>% 
  leaflet() %>% 
  setView(lng=0,lat=0,zoom= 1) %>% 
  addTiles()  %>% 
  addPolygons(fillColor = ~pal(Perceived_Threat), weight=1,color = "#BDBDC3"
              ,fillOpacity=0.9 ,
             popup = popUp) %>% 

  addLegend(colors = c(rev(RColorBrewer::brewer.pal(theBins, "RdYlGn")), "#808080"),  
            bins = theBins, 
            position = 'bottomleft', 
            title = "Ranking Band", 
            labels = labs) %>% 
  mapOptions(zoomToLimits="first") 
  }



theMap
})

leafletOutput("map")
```



Some interesting variation here. Europe and North America were among the most aware but South American countries seemed more keen to pin the blame on human actions, a view to which 97% of climate scientists apparently concur

---


Here are some charts comparing the variables. Hover over point for Country

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}


data <- cbind(data, id = seq_len(nrow(data)))

 
 all_values<- function(x) {
   if(is.null(x)) return(NULL)
   row <- data[data$id == x$id,c("Country")]
   paste0( format(row), collapse = "<br />")
 }


 
data %>% 
  ggvis(~Awareness,~Caused_by_Humans,key := ~id) %>%
   layer_points() %>%
   add_tooltip(all_values, "hover") %>% 
   add_axis("y", title="Caused by Humans")


data %>% 
  ggvis(~Awareness,~Perceived_Threat,key := ~id) %>%
   layer_points() %>%
   add_tooltip(all_values, "hover")  %>% 
   add_axis("y", title="Perceived Threat")

data %>% 
  ggvis(~Perceived_Threat,~Caused_by_Humans,key := ~id) %>%
   layer_points() %>%
   add_tooltip(all_values, "hover") %>% 
   add_axis("y", title="Caused by Humans")%>% 
   add_axis("x", title="Perceived Threat")


```

